\mytitle

### 2015-08-05 MSc_diss.tex
Checked numbers of reactions and figure for the first time

\begin{figure}[t] % O3 budget
\caption{Summary of tropospheric ozone global budget model and observation estimates for present (about 2000) conditions. All uncertainties quoted as 1 standard deviation (68\% confidence interval). Adapted from \citep{Myhre2013}}
\centering
\label{tab:O3budget}
\includegraphics[width=0.7\linewidth]{o3_studies.eps}
\end{figure}

@@

\subsection{Alkyl nitrates}\label{sec:intro_AN}
Alkyl nitrates ($RONO_2$) are organic trace gases, which through one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}. This mechanism is present in the following section.
\subsubsection*{Alkyl nitrates chemistry}\label{into_ANchem}
The photochemical formation of the alkyl nitrates starts when chemistry of the troposphere reaches the regime of $O_3$ production, i.e. when reactions of the peroxy radicals with $NO$ start to compete with their self- and cross-reactions more effectively due to an increase in $NO_x$ mixing ratio. In this case the reaction between the alkyl peroxy radical ($RO_2$, where $R$ denotes methyl group, $CH_3$) and $NO$ can reveal its two-channel nature \citep{Day2003}:
\begin{subequations} \label{reac:RONO2alpha}
\begin{align}
RO_2 + NO &\xrightarrow{1-\alpha_3} RO + NO_2 \label{reac:RO+NO=aphaRO+NO2}\\
RO_2 + NO &\xrightarrow{\alpha_3} RONO_2 \label{reac:RO+NO=aphaRONO2}
\end{align}
\end{subequations}
The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle (Figure ???), and as \citep{Roberts1998} estimated it is likely to begin to occur when the $NO$ mixing ratio increases up to 100-200 ppt. The yield of alkyl nitrate formation is controlled by the branching ratio ($\alpha_3$), which is temperature and pressure dependent and increases with the number of carbons in the alkyl peroxy radical \citep{Roberts1990}.

An alternative photochemical pathway of the formation of alkyl nitrates is the reaction of the alkoxy radical ($RO$) with $NO_2$:
\begin{equation}\label{reac:RO+NO2=RONO2}
RO + NO_2 \rightarrow RONO_2
\end{equation}
This pathway has been shown to become dominant over reaction (\ref{reac:RO+NO=aphaRONO2}) only for the production of lower (methyl ($C_1$), ethyl ($C_2$)) nitrates when $NO_x$ mixing ratio exceeds 35 ppb \citep{Archibald2007}, i.e. in highly polluted environments.

@@

The lifetimes of alkyl nitrates depend on the seasonality of the photolysis and $OH$ sink, which in turn vary with latitude and altitude. According to \citep{Clemitshaw1997} and \citep{Turberg1990} alkyl nitrates have atmospheric lifetimes of several days to several years, which makes them temporary reservoirs for the short lived $NO_x$ compounds. Being transported from polluted areas to remote background regions and eventually destroyed alkyl nitrates can 'release' $NO_2$ back to the atmosphere, thereby contributing to tropospheric ozone formation.
\subsection{Ozone, alkyl nitrates and carbon bonds}\label{sec:intro_CB}
In the previous sections we covered the chemistry that controls the production of $O_3$ in the troposphere. A traditional way to assess the efficiency of this production is to calculate the ratio:
\begin{matriz}\label{eq:OPE}
OPE = \dfrac{k_{HO_2+NO}[HO_2][NO]}{k_{OH+NO_2}[OH][NO_2]} = \dfrac{P_{O_3}}{L_{NO_x}}
\end{matriz}
where $OPE$ is the ozone production efficiency, $k_{HO_2+NO}$ - the rate coefficient of reaction (\ref{reac:HO2+NO=NO2+OH}), $k_{OH+NO_2}$ - the rate coefficient of reaction (\ref{reac:OH+NO2=HNO3}), $[X]$ - mixing ratios of corresponding species, $P_{O_3}$ – the $O_3$ production rate, $L_{NO_x}$ – the $NO_x$ loss rate. $OPE$ is determined by the amount of $HO_2$ reacting with $NO$ to form $NO_2$ and $OH$ (i.e., $k_{HO_2+NO}[HO_2][NO]$, (\ref{reac:HO2+NO=NO2+OH})) and the amount of $NO_x$ removed from the system by reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$, reaction (\ref{reac:OH+NO2=HNO3})). In other words, $OPE$ shows the amount of ozone created ($P_{O_3}$) per molecule of $NO_x$ removed ($L_{NO_x}$) from the system \citep{Jacob1999}. Due to the reasons described in section \ref{sec:chem_mech} (comment at the bottom of the page!"No representation of the nitric acid ($HNO_3$) in the model"), we use another expression for calculations of the $O_3$ production efficiency. This expression has been recently proposed by Professor Matthew Evans (University of York) amid the mounting concerns about the need of thorough intercomparison of chemical mechanisms of global atmospheric models (including those used in the IPCC reports), why it is more universal and applicable to any kind of model.

@@

Temporal changes in reactive nitrogen partitioning in the DWH plume (Figure \ref{fig:Neuman2012}) nicely illustrate that in fresh part of the plume (Figure ??? a) $NO_x$ accounted for over 80\% of plume $NO_y$, in aged plumes (Figure ??? b), $NO_x$ oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight $C_5$–$C_{10}$ $RONO_2$ accounted for only 22\% of plume $NO_y$. It means that $C_6$–$C_{10}$ $RONO_2$ account for 78\% of the aged plume $NO_y$\citep{Neuman2012}. This result differs dramatically from previous studies, where the estimated contribution of $RONO_2$ to $NO_y$ was much lower: approximately 4\% in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003], approx. 10\% in rural areas [Parrish et al., 1993], a few percent in urban air masses [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010].

Probably the most detailed investigation of the formation processes and photochemical evolution of alkyl nitrates (but not their impact on $O_3$) has been done by \citep{Sommariva2008}. Using a subset of the Master Chemical Mechanism v3.1 (MCM) they constructed a highly detailed chemical model and run it for 5 days under two scenarios (Table \ref{tab:Sommariva2008init}) to explore the sensitivity of the evolution of the chemical composition of an air mass emitted from an urban environment to the initial conditions. As it seen in Figure \ref{fig:Sommariva2008fig} according to the modelling results $O_3$ is produced and $NO_x$ is removed from the system with the consequent formation of $HNO_3$ via the reaction of $NO_2$ with $OH$, which broadly agrees with the observations [(Neuman et al., 2006).]. In scenario A the concentration of $NO_x$ after the first day is about 0.5–1 ppb for $NO_2$ and approximately 100 ppt for $NO$, which is high enough to sustain the production of $O_3$ to up to 135 ppb and the formation of alkyl nitrates during the simulation. In scenario B (Figure \ref{fig:Sommariva2008fig} right) the increase in $O_3$ is slower (note that initial concentration is 0) and the maximum concentration after five days is about 100 ppb. The concentration of $NO_x$ remains relatively high during the second (10–15 ppb) and third day (3–5 ppb) and then decreases to $<$200 ppt of $NO$ and $<$1ppb of $NO_2$ on the fourth and fifth days, therefore allowing the formation of alkyl nitrates too. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}\label{tab:Sommariva2008init}
\caption{Initial values of the some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}
\centering
\begin{tabular}{ccc}
\hline
Parameter & Initial value Scenario A & Scenario B \\
\hline
$CO$      & 332.5 & 232.5 \\
$NO_x$    & 34.9  & 46.5  \\
$O_3$     & 90    & 0   \\
\hline
\end{tabular}
\end{table}
\begin{figure}\label{fig:Sommariva2008fig}
\includegraphics[width=\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Sommariva2008_Fig.1}.png}
\caption{Modelled concentrations of $O_3$, $NO_x$ and $HNO_3$ for scenario A (left) and B (right). Source: \citep{Sommariva2008}.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Comparison of modelling results with the values observed inside a typical urban plume sampled using the NOAA WP-3D aircraft from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign enabled \citep{Sommariva2008} to conclude that the chemistry of alkyl nitrates is well represented in the MCM model. They found that short-chain $<C_4$ alkyl nitrates can be formed from several precursors, while long-chain ($C_5$) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane. They also found that at lower $O_3$ and $CO$ initial concentrations and higher $NOx$ (scenario A) the importance of the reactions between alkane and $OH$ is typically lower during the first two-three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important.

The first 3-D chemical transport  model simulations examining the impact of methyl and ethyl nitrates on tropospheric ozone chemistry in remote regions have been conducted by \citep{Neu2008}. In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM [Carver et al., 1997] and by matching their observed atmospheric abundances derived a total oceanic emission flux of these two nitrates (0.35 Tg of N per year). Comparing two simulations with and without spatially uniform emissions of methyl and ethyl nitrates they found that the largest impact of oceanic alkyl nitrates is seen over the equatorial Western Pacific (Figure ???). Atmospheric circulation (convergence of the trade winds) in this region isolates the marine boundary layer, resulting in sustained photochemical loss of $O_3$ (Figure ??? c) and at the same time accumulation of alkyl nitrates (Figure ??? a), which drive up boundary layer $NO_x$ abundances by as much as 250\% (Figure ??? b). Such combination of factors leads to increases in $O_3$ of up to 20\% (Figures ??? c and d), with the largest increases occurring from May to October.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t!]\label{fig:Neu2008}
	\centering
	\begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.2_MeONO2_OCT_merge}.png}
        \caption{}
    \end{subfigure}%
    ~ % glues subplots on the same line somehow...
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3contol}.png}
        \caption{}
    \end{subfigure}
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_NOxdiff}.png}
        \caption{}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3diff}.png}
        \caption{}
    \end{subfigure}
\caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (c) $O_3$, and percent difference (perturbation - control) in (b) $NO_x$, (d) $O_3$. All averaged over the boundary layer (below 850 mb). Adapted from \citep{Neu2008}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The full list of reactions can be found in Appendix ?? \ref{sec:appendix1}. It is important to mention that due to purely theoretical purpose of this research and its focus on investigation of the impact of alkyl nitrate chemistry on ozone production some of the reactions that lead to formation of species that serve as a sink for $NO_x$ are not included into the chemical mechanism. However, these reactions play an important role in the chemistry of the troposphere, and below here is a description of their role and anticipated effect of their exclusion.

@@

\begin{table} % model setup: VOCs
\caption{Model setup: VOCs concentrations (ppb)}\label{tab:setupVOCs}
\centering
\begin{tabular}{ccccccccc}
\hline
Experiment number & $CH_4$ & $C_2H_6$ & $C_3H_8$ & $nC_4H_{10}$ & $iC_4H_{10}$ & $nC_5H_{12}$ & $iC_5H_{12}$ \\
\hline
A  & 1780 &	0.620 &	0.100 &	0.050 &	0.250 &	0.250 &	0.250 \\
B  & 1807 &	1.258 &	0.390 &	0.245 &	0.325 &	0.275 &	0.425 \\
C  & 1834 &	1.896 &	0.680 &	0.440 &	0.400 &	0.300 &	0.600 \\
D  & 1861 &	2.534 &	0.970 &	0.635 &	0.475 &	0.325 &	0.775 \\
E  & 1888 &	3.172 &	1.260 &	0.830 &	0.550 &	0.350 &	0.950 \\
F  & 1915 &	3.810 &	1.550 &	1.025 &	0.625 &	0.375 &	1.125 \\
G  & 1942 &	4.448 &	1.840 &	1.220 &	0.700 &	0.400 &	1.300 \\
H  & 1969 &	5.086 &	2.130 &	1.415 &	0.775 &	0.425 &	1.475 \\
I  & 1996 &	5.724 &	2.420 &	1.610 &	0.850 &	0.450 &	1.650 \\
J  & 2023 &	6.362 &	2.710 &	1.805 &	0.925 &	0.475 &	1.825 \\
K  & 2050 &	7.000 &	3.000 &	2.000 &	1.000 &	0.500 &	2.000 \\
\hline
\end{tabular}
\end{table}
\subsection{Model performance}
Due to imperfection of the chemical mechanism time evolution of species concentrations given by the model does not completely agree with the photochemical theory, but reproduces it to some extent. Figure \ref{fig:model_validation} shows an example of output from the model, namely the changes in mixing ratios of the main species within 24-hours of simulation time under contrasting $NO_x$ and $VOCs$ conditions. The values shown in Figure \ref{fig:model_validation} have been justified from the theoretical point of view and also compared with those derived from the Atmospheric Chemistry Learner's Box Model (ACLBM, \href{http://guille.uea.ac.uk/ACLBM/}{http://guille.uea.ac.uk/ACLBM/}). ACLBM is an on-line box model for atmospheric chemistry developed by Prof. Roland von Glasow (University of East Anglia) that allows the selection of several chemical reaction mechanisms, choice of initial concentrations and emission fluxes. For comparison with our modelling data the ACLBM was set up and initialized with values from Tables \ref{tab:setupO3CO}, \ref{tab:setupNOx} and \ref{tab:setupVOCs}, and the '$O_3$-$NO_x$-$CO$-$CH_4$' chemical mechanism was chosen as it is the closest version of the chemical mechanism available in the ACLBM to our model.

>>>>

\begin{figure}[t] % O3 budget
\centering
\includegraphics[width=0.7\linewidth]{o3_studies.eps}
\caption{Summary of tropospheric ozone global budget model and observation estimates for present (about 2000) conditions. All uncertainties quoted as 1 standard deviation (68\% confidence interval). Adapted from \citep{Myhre2013}}\label{fig:O3budget}
\end{figure}

@@

\subsection{Alkyl nitrates}\label{sec:intro_AN}
Alkyl nitrates ($RONO_2$) are organic trace gases, which through one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}. This mechanism is present in the following section.
\subsubsection*{Alkyl nitrates chemistry}\label{into_ANchem}
The photochemical formation of the alkyl nitrates starts when chemistry of the troposphere reaches the regime of $O_3$ production, i.e. when reactions of the peroxy radicals with $NO$ start to compete with their self- and cross-reactions more effectively due to an increase in $NO_x$ mixing ratio. In this case the reaction between the alkyl peroxy radical ($RO_2$, where $R$ denotes methyl group, $CH_3$) and $NO$ can reveal its two-channel nature \citep{Day2003}:
\begin{subequations} \label{reac:RONO2alpha}
\begin{align}
RO_2 + NO &\xrightarrow{1-\alpha_3} RO + NO_2 \label{reac:RO2+NO=aphaRO+NO2}\\
RO_2 + NO &\xrightarrow{\alpha_3} RONO_2 \label{reac:RO2+NO=aphaRONO2}
\end{align}
\end{subequations}
The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle (Figure \ref{fig:O3scheme}), and as \citep{Roberts1998} estimated it is likely to begin to occur when the $NO$ mixing ratio increases up to 100-200 ppt. The yield of alkyl nitrate formation is controlled by the branching ratio ($\alpha_3$), which is temperature and pressure dependent and increases with the number of carbons in the alkyl peroxy radical \citep{Roberts1990}.

An alternative photochemical pathway of the formation of alkyl nitrates is the reaction of the alkoxy radical ($RO$) with $NO_2$:
\begin{equation}\label{reac:RO+NO2=RONO2}
RO + NO_2 \rightarrow RONO_2
\end{equation}
This pathway has been shown to become dominant over reaction (\ref{reac:RO2+NO=aphaRONO2}) only for the production of lower (methyl ($C_1$), ethyl ($C_2$)) nitrates when $NO_x$ mixing ratio exceeds 35 ppb \citep{Archibald2007}, i.e. in highly polluted environments.

@@

The lifetimes of alkyl nitrates depend on the seasonality of the photolysis and $OH$ sink, which in turn vary with latitude and altitude. According to \citep{Clemitshaw1997} and \citep{Turberg1990} alkyl nitrates have atmospheric lifetimes of several days to several years, which makes them temporary reservoirs for the short lived $NO_x$ compounds. Being transported from polluted areas to remote background regions and eventually destroyed alkyl nitrates can 'release' $NO_2$ back to the atmosphere, thereby contributing to tropospheric ozone formation.
\subsection{Ozone, alkyl nitrates and carbon bonds}\label{sec:intro_CB}
In the previous sections we covered the chemistry that controls the production of $O_3$ in the troposphere. A traditional way to assess the efficiency of this production is to calculate the ratio:
\begin{matriz}\label{eq:OPE}
OPE = \dfrac{k_{HO_2+NO}[HO_2][NO]}{k_{OH+NO_2}[OH][NO_2]} = \dfrac{P_{O_3}}{L_{NO_x}}
\end{matriz}
where $OPE$ is the ozone production efficiency, $k_{HO_2+NO}$ - the rate coefficient of reaction (\ref{reac:HO2+NO=NO2+OH}), $k_{OH+NO_2}$ - the rate coefficient of reaction (\ref{reac:OH+NO2=HNO3}), $[X]$ - mixing ratios of corresponding species, $P_{O_3}$ – the $O_3$ production rate, $L_{NO_x}$ – the $NO_x$ loss rate. $OPE$ is determined by the amount of $HO_2$ reacting with $NO$ to form $NO_2$ and $OH$ (i.e., $k_{HO_2+NO}[HO_2][NO]$, (\ref{reac:HO2+NO=NO2+OH})) and the amount of $NO_x$ removed from the system by reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$, reaction (\ref{reac:OH+NO2=HNO3})). In other words, $OPE$ shows the amount of ozone created ($P_{O_3}$) per molecule of $NO_x$ removed ($L_{NO_x}$) from the system \citep{Jacob1999}. Due to the reasons described in section \ref{sec:method_chem} \footnote{No representation of the nitric acid ($HNO_3$) in the model.}, we use another expression for calculations of the $O_3$ production efficiency. This expression has been recently proposed by Professor Matthew Evans (University of York) amid the mounting concerns about the need of thorough intercomparison of chemical mechanisms of global atmospheric models (including those used in the IPCC reports), why it is more universal and applicable to any kind of model.

@@

Temporal changes in reactive nitrogen partitioning in the DWH plume (Figure \ref{fig:Neuman2012}) nicely illustrate that in fresh part of the plume (Figure \ref{fig:Neuman2012}a) $NO_x$ accounted for over 80\% of plume $NO_y$, in aged plumes (Figure \ref{fig:Neuman2012}b), $NO_x$ oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight $C_5$–$C_{10}$ $RONO_2$ accounted for only 22\% of plume $NO_y$. It means that $C_6$–$C_{10}$ $RONO_2$ account for 78\% of the aged plume $NO_y$ \citep{Neuman2012}. This result differs dramatically from previous studies, where the estimated contribution of $RONO_2$ to $NO_y$ was much lower: approximately 4\% in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003], approx. 10\% in rural areas [Parrish et al., 1993], a few percent in urban air masses [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010].

Probably the most detailed investigation of the formation processes and photochemical evolution of alkyl nitrates (but not their impact on $O_3$) has been done by \citep{Sommariva2008}. Using a subset of the Master Chemical Mechanism v3.1 (MCM) they constructed a highly detailed chemical model and run it for 5 days under two scenarios (Table \ref{tab:Sommariva2008init}) to explore the sensitivity of the evolution of the chemical composition of an air mass emitted from an urban environment to the initial conditions. As it seen in Figure \ref{fig:Sommariva2008fig} according to the modelling results $O_3$ is produced and $NO_x$ is removed from the system with the consequent formation of $HNO_3$ via the reaction of $NO_2$ with $OH$, which broadly agrees with the observations [(Neuman et al., 2006).]. In scenario A the concentration of $NO_x$ after the first day is about 0.5–1 ppb for $NO_2$ and approximately 100 ppt for $NO$, which is high enough to sustain the production of $O_3$ to up to 135 ppb and the formation of alkyl nitrates during the simulation. In scenario B (Figure \ref{fig:Sommariva2008fig} right) the increase in $O_3$ is slower (note that initial concentration is 0) and the maximum concentration after five days is about 100 ppb. The concentration of $NO_x$ remains relatively high during the second (10–15 ppb) and third day (3–5 ppb) and then decreases to $<$200 ppt of $NO$ and $<$1ppb of $NO_2$ on the fourth and fifth days, therefore allowing the formation of alkyl nitrates too. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\caption{Initial values of the some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}\label{tab:Sommariva2008init}
\centering
\begin{tabular}{ccc}
\hline
Parameter & Initial value Scenario A & Scenario B \\
\hline
$CO$      & 332.5 & 232.5 \\
$NO_x$    & 34.9  & 46.5  \\
$O_3$     & 90    & 0   \\
\hline
\end{tabular}
\end{table}
\begin{figure}
\includegraphics[width=\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Sommariva2008_Fig.1}.png}
\caption{Modelled concentrations of $O_3$, $NO_x$ and $HNO_3$ for scenario A (left) and B (right). Source: \citep{Sommariva2008}.}\label{fig:Sommariva2008fig}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Comparison of modelling results with the values observed inside a typical urban plume sampled using the NOAA WP-3D aircraft from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign enabled \citep{Sommariva2008} to conclude that the chemistry of alkyl nitrates is well represented in the MCM model. They found that short-chain $<C_4$ alkyl nitrates can be formed from several precursors, while long-chain ($C_5$) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane. They also found that at lower $O_3$ and $CO$ initial concentrations and higher $NOx$ (scenario A) the importance of the reactions between alkane and $OH$ is typically lower during the first two-three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important.

The first 3-D chemical transport  model simulations examining the impact of methyl and ethyl nitrates on tropospheric ozone chemistry in remote regions have been conducted by \citep{Neu2008}. In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM [Carver et al., 1997] and by matching their observed atmospheric abundances derived a total oceanic emission flux of these two nitrates (0.35 Tg of N per year). Comparing two simulations with and without spatially uniform emissions of methyl and ethyl nitrates they found that the largest impact of oceanic alkyl nitrates is seen over the equatorial Western Pacific (Figure \ref{fig:Neu2008}). Atmospheric circulation (convergence of the trade winds) in this region isolates the marine boundary layer, resulting in sustained photochemical loss of $O_3$ (Figure \ref{fig:Neu2008M_O3contol}) and at the same time accumulation of alkyl nitrates (Figure \ref{fig:Neu2008M_MeONO2}), which drive up boundary layer $NO_x$ abundances by as much as 250\% (Figure \ref{fig:Neu2008M_NOxdiff}). Such combination of factors leads to increases in $O_3$ of up to 20\% (Figures \ref{fig:Neu2008M_O3contol} and \ref{fig:Neu2008M_O3diff}), with the largest increases occurring from May to October.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t!]
	\centering
	\begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.2_MeONO2_OCT_merge}.png}
        \caption{}\label{fig:Neu2008M_MeONO2}
    \end{subfigure}%
    ~ % glues subplots on the same line somehow...
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3contol}.png}
        \caption{}\label{fig:Neu2008M_O3contol}
    \end{subfigure}
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_NOxdiff}.png}
        \caption{}\label{fig:Neu2008M_NOxdiff}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3diff}.png}
        \caption{}\label{fig:Neu2008M_O3diff}
    \end{subfigure}
\caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (b) $O_3$, and percent difference (perturbation - control) in (c) $NO_x$, (d) $O_3$. All averaged over the boundary layer (below 850 mb). Adapted from \citep{Neu2008}}\label{fig:Neu2008}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The full list of reactions can be found in Appendix \ref{sec:appendix1}. It is important to mention that due to purely theoretical purpose of this research and its focus on investigation of the impact of alkyl nitrate chemistry on ozone production some of the reactions that lead to formation of species that serve as a sink for $NO_x$ are not included into the chemical mechanism. However, these reactions play an important role in the chemistry of the troposphere, and below here is a description of their role and anticipated effect of their exclusion.

@@

\begin{table} % model setup: VOCs
\caption{Model setup: VOCs concentrations (ppb)}\label{tab:setupVOCs}
\centering
\begin{tabular}{ccccccccc}
\hline
Experiment number & $CH_4$ & $C_2H_6$ & $C_3H_8$ & $nC_4H_{10}$ & $iC_4H_{10}$ & $nC_5H_{12}$ & $iC_5H_{12}$ \\
\hline
A  & 1780 &	0.620 &	0.100 &	0.050 &	0.250 &	0.250 &	0.250 \\
B  & 1807 &	1.258 &	0.390 &	0.245 &	0.325 &	0.275 &	0.425 \\
C  & 1834 &	1.896 &	0.680 &	0.440 &	0.400 &	0.300 &	0.600 \\
D  & 1861 &	2.534 &	0.970 &	0.635 &	0.475 &	0.325 &	0.775 \\
E  & 1888 &	3.172 &	1.260 &	0.830 &	0.550 &	0.350 &	0.950 \\
F  & 1915 &	3.810 &	1.550 &	1.025 &	0.625 &	0.375 &	1.125 \\
G  & 1942 &	4.448 &	1.840 &	1.220 &	0.700 &	0.400 &	1.300 \\
H  & 1969 &	5.086 &	2.130 &	1.415 &	0.775 &	0.425 &	1.475 \\
I  & 1996 &	5.724 &	2.420 &	1.610 &	0.850 &	0.450 &	1.650 \\
J  & 2023 &	6.362 &	2.710 &	1.805 &	0.925 &	0.475 &	1.825 \\
K  & 2050 &	7.000 &	3.000 &	2.000 &	1.000 &	0.500 &	2.000 \\
\hline
\end{tabular}
\end{table}

\subsection{Model performance}
Due to imperfection of the chemical mechanism time evolution of species concentrations given by the model does not completely agree with the photochemical theory, but reproduces it to some extent. Figure \ref{fig:model_validation} shows an example of output from the model, namely the changes in mixing ratios of the main species within 24-hours of simulation time under contrasting $NO_x$ and $VOCs$ conditions. The values shown in Figure \ref{fig:model_validation} have been justified from the theoretical point of view and also compared with those derived from the Atmospheric Chemistry Learner's Box Model (ACLBM, \href{http://guille.uea.ac.uk/ACLBM/}{http://guille.uea.ac.uk/ACLBM/}). ACLBM is an on-line box model for atmospheric chemistry developed by Prof. Roland von Glasow (University of East Anglia) that allows the selection of several chemical reaction mechanisms, choice of initial concentrations and emission fluxes. For comparison with our modelling data the ACLBM was set up and initialized with values from Tables \ref{tab:setupO3CO}, \ref{tab:setupNOx} and \ref{tab:setupVOCs}, and the '$O_3$-$NO_x$-$CO$-$CH_4$' chemical mechanism was chosen as it is the closest version of the chemical mechanism available in the ACLBM to our model.

The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere, and theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: at low $NO_x$ levels the regime is characterized by net $O_3$ removal, at intermediate $NO_x$ (also called $NO_x$-sensitive) - by net $O_3$ formation, and at high $NO_x$ - by net $O_3$ formation or removal depending more on $VOCs$ concentration (therefore called $VOC$-sensitive) (e.g., \citep{Fowler2008}). When the prevailing $NO_x$ levels are low, a net $O_3$ removal takes place because there is not enough $NO$ around to react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$, which photolysis in conjunction with the reaction (\ref{reac:O3P+O2=O3}) is the main source of $O_3$ in the troposphere. Some additional $O_3$ removal in this regime occurs due to reaction of $HO_2$ radicals with $O_3$ leading to the regeneration of $OH$ radicals as part of an $O_3$ depleting $OH$-$HO_2$ ($HO_x$) cycle. In other words, the following reaction sequence predominates:
\begin{equation}
O_3 + hv \rightarrow O(^1D) + O_2\tag{\ref{reac:O3+hv=O1D+O2}}
\end{equation}
\begin{equation}
O(^1D) + H_2O \rightarrow OH + OH\tag{\ref{reac:O1D+H2O=OH+OH}}
\end{equation}
\begin{equation}
O(^1D) \xrightarrow{M} O(^3P)\tag{\ref{reac:O1D=O3P}}
\end{equation}
\begin{equation}
O(^3P) + O_2 \xrightarrow{M} O_3\tag{\ref{reac:O3P+O2=O3}}
\end{equation}
\begin{equation}
OH + CO \rightarrow HO_2 + CO_2\tag{\ref{reac:OH+CO=H+CO2}}
\end{equation}
\begin{equation}\label{reac:OH+RH=RO2+H2O}
OH + RH \rightarrow RO_2 + H_2O
\end{equation}
\begin{equation}
HO_2 + HO_2 \rightarrow H_2O_2 + O_2\tag{\ref{reac:HO2+HO2=H2O2+O2}}
\end{equation}
\begin{equation}\label{reac:RO2+HO2=ROOH+O2}
RO_2 + HO_2 \rightarrow ROOH + O_2
\end{equation}
\begin{equation}
O_3 + HO_2 \rightarrow OH + 2O_2\tag{\ref{reac:HO2+O3=OH+2O2}}
\end{equation}
At low $NO_x$ regime (for example, 50 ppt) our modelling data agrees well with the theory (Figure \ref{fig:O3valid}). $O_3$ mixing ratio decreases with time, and this decrease is comparable to one derived from ACLBM (by 6 ppb and 4.5 ppb at low and high $VOCs$ concentrations respectively). $NO$ is rapidly (almost instantly) converted to $NO_2$ representing a state equilibrium between the two molecules in the troposphere (Figures \ref{fig:NOvalid}, \ref{fig:NO2valid}). $CO$ is lost due to its reaction with $OH$ (Figure \ref{fig:COvalid}). Mixing ratios of other species in our model are of the order of magnitude as in ACLBM.
